# Module and event timing

A next step could be to understand how long time is being spent in particular _art_ modules and on a more event by event level. Because the bulk of the simulation runs in the `artg4` producer, this measurement may be less useful compared to a many module _art_ job. But it will still be instructive so it is good to do. 

The `TimeTracker` _art_ service provides timing information about each module and statistics over events. It can also create a database with event by event timing information. You can run `TimeTracker` by adding the following to your FCL file...

```
services : {

// ...

  TimeTracker : {
    printSummary : true
    dbOutput : {
       filename : "v7_timing.db"
       overwrite : true
    }
  }
  
// ...

}
```

See http://cdcvs.fnal.gov/redmine/projects/art/wiki/TimeTracker for more information and instructions. 

## Summary printout

The default usage of the service and the configuration described here prints out a summary of timing information at the end of the job. For our example comparing `gm2 v6_04_00` and `gm2 v7_03_00`, we get the following output...

```
# v6_04_00
TimeReport ---------- Time  Summary ---[sec]----
TimeReport CPU = 48.492349 Real = 48.291652

================================================================================================================================
TimeTracker printout (sec)                        Min           Avg           Max         Median          RMS         nEvts
================================================================================================================================
Full event                                    0.00430586     0.0969926     0.979898      0.0567863     0.111273        500
--------------------------------------------------------------------------------------------------------------------------------
path1:randomsaver:RandomNumberSaver           1.5818e-05    2.97016e-05   0.000218545   2.39715e-05   1.6956e-05       500
path1:artg4:artg4Main                         0.00312922     0.0931658     0.970814      0.0553228     0.108083        500
path1:TriggerResults:TriggerResultInserter     9.438e-06    1.54518e-05   9.9472e-05    1.2985e-05    8.65874e-06      500
end_path:out1:RootOutput                      0.000232495   0.00371359     0.163321     0.00142215    0.00797466       500
================================================================================================================================
```

and

```
# v7_03_00
TimeReport ---------- Time  Summary ---[sec]----
TimeReport CPU = 457.772410 Real = 259.748750

================================================================================================================================
TimeTracker printout (sec)                        Min           Avg           Max         Median          RMS         nEvts
================================================================================================================================
Full event                                     0.0103769     0.519505       4.16775      0.257809      0.594227        500
--------------------------------------------------------------------------------------------------------------------------------
path1:randomsaver:RandomNumberSaver           3.2923e-05    5.95679e-05   0.00119983    4.83085e-05   5.79987e-05      500
path1:artg4:artg4Main                           0.01027      0.519357       4.16762      0.257689      0.594228        500
path1:TriggerResults:TriggerResultInserter    1.7502e-05    2.29663e-05   9.7601e-05    2.09805e-05   7.6282e-06       500
end_path:out1:RootOutput                       2.405e-06    3.71508e-06   2.0902e-05    3.5705e-06    1.11764e-06      500
end_path:out1:RootOutput(write)               0.000217776   0.00289537     0.0216031    0.00107864     0.0031938       500
================================================================================================================================
```

Note the statistics for events. The only producer that really matters here is `artg4::artg4Main`, which runs Geant. On average, for `v6` we see it took Geant 0.093s/muon and for `v7` we see it took Geant 0.52s/muon. This is a dramatic difference. It is also interesting to look at the extremes - there was an event run by `v7` that took over four seconds!

## Database 

As shown in the FCL fragment above, `TimeTracker` can optionally output a database with timing information for each event. Let's look at the timing information for `v7`. For the analysis we'll use _R_. 

```{r}
library(tidyverse)

db <- src_sqlite("data/v7_timing.db")
db
```

The `TimeEvent` table has the total time for each event. 
```{r}
te <- tbl(db, "TimeEvent")
te %>% collect()
```

The `TimeModule` table has the time for each event broken down by module.
```{r}
tm <- tbl(db, "TimeModule")
tm %>% collect()
```

The `TimeReport` table gives the same information as the printed output from the `TimeTracker` service. 
```{r}
tr <- tbl(db,"TimeReport")
tr %>% collect()
```

We can make plots... 
```{r}
tm %>% filter(PathModuleId == 'path1:artg4:artg4Main') %>% collect() -> artg4Time 
artg4Time %>% ggplot( aes(x=Event, y=Time) ) + geom_point() 
```

```{r}
artg4Time %>% ggplot( aes(x=Time) ) + geom_histogram(bins=50)
```

